
%%%%%%%%%%%%%  Start of GENERATED TEXT  %%%%%%%%%%%%\n

\newpage
\setcounter{page}{1}
\setcounter{section}{0}
\setcounter{equation}{0}
\setcounter{figure}{0}
\setcounter{table}{0}
\setcounter{footnote}{0}
\renewcommand{\ehkol}{Kolomiiets B. Yu., Karpliuk Ye. S.}
\renewcommand{\ohkol}{Improving EMG Signal Classification with Transfer Learning Under Low-Data and Cross-Subject Conditions}
\Vykh{radap20NN:FirstPage}{radap20NN:LastPage}
\begin{multicols}{2}
[
\UDC{}
\renewcommand{\ehkol}{Kolomiiets B. Yu., Karpliuk Ye. S.}
\renewcommand{\ohkol}{Improving EMG Signal Classification with Transfer Learning Under Low-Data and Cross-Subject Conditions}
\Title{Improving EMG Signal Classification with Transfer Learning Under Low-Data and Cross-Subject Conditions}\label{radap20NN:FirstPage}
\Authors{Kolomiiets B. Yu., Karpliuk Ye. S.}
\aff{National Technical University of Ukraine "Igor Sikorsky Kyiv Polytechnic Institute", Kyiv, Ukraine}
\Address{}{kolomiiets.bohdan.yuriyovych@gmail.com}
\AbsKeywords{
Surface electromyography (sEMG) is a non-invasive method used for human–machine interaction, but its practical use is often limited by large differences between individuals and the effort needed to train models for each new user. This study explores whether transfer learning can help solve these problems when using deep learning to classify hand and wrist gestures. The experiments were conducted on a dataset that includes eleven gestures, each repeated eight times by 22 healthy participants. Three training approaches were tested: (i) training and testing on the same subject (intra-subject), (ii) training on some subjects and testing on a new one (inter-subject), and (iii) transfer learning with and without resetting the fully connected (FC) output layer of the convolutional neural network (CNN). All models were evaluated using a leave-one-out cross-validation strategy across both subjects and repetitions.
Results show that both transfer learning methods performed better than the other two approaches in terms of classification accuracy. The best result was achieved when the FC layer was reset before fine-tuning (F1-score = 0.907, $\sigma = 0.074$). Wilcoxon signed-rank statistical tests confirmed that these improvements were statistically significant, even when only a few repetitions were used for fine-tuning. In fact, using transfer learning with just four repetitions instead of eight gave similar accuracy to training from scratch on eight repetitions.
These findings suggest that fine-tuning pre-trained models can significantly lower the effort required to adapt EMG-based systems for new users, offering a practical and effective approach for developing practical interfaces suited to assistive and rehabilitation applications.
}
{Keywords}{
surface electromyography (sEMG), gesture recognition, transfer learning, deep learning, convolutional neural networks (CNN), inter-subject variability, model generalization, fine-tuning strategies, subject adaptation, myoelectric control, EMG-based interface, cross-validation, biomedical signal processing, rehabilitation technologies, low-effort calibration
 }
{\href{http://radap.kpi.ua/radiotechnique/article/view/20NN}{10.20535/RADAP.2025.\#\#.\pageref{radap20NN:FirstPage}-\pageref{radap20NN:LastPage}}}
]

\input{radap20NN}  

\TitleSecond{Покращення класифікації ЕМГ-сигналів за допомогою трансферного навчання в умовах обмежених даних та варіативності між користувачами}
\Auth{Коломієць Б. Ю., Карплюк Є. С.}
\AbsK{
Поверхнева електроміографія (sEMG) це неінвазивний метод взаємодії людини з машиною, проте її практичне застосування часто обмежується значною варіативністю між користувачами та необхідністю налаштування моделей під кожного нового користувача. У цьому дослідженні розглядається, чи може трансферне навчання допомогти подолати ці труднощі під час класифікації жестів кисті та зап’ястя із застосуванням методів глибокого навчання. Експерименти проводилися на датасеті, що містить одинадцять жестів, кожен з яких повторено вісім разів 22 здоровими учасниками. Було протестовано три підходи до навчання: (i) навчання та тестування на даних одного учасника (внутрішньоособове навчання), (ii) навчання на даних декількох учасників і тестування на новому (міжособове узагальнення), (iii) трансферне навчання з донавчанням, із скиданняv та без скидання повнозв’язного (FC) вихідного шару згорткової нейронної мережі (CNN). Усі моделі оцінювались за допомогою стратегії перехресної перевірки з виключенням одного елемента (leave-one-out) як по учасниках, так і по повтореннях.
Результати показали, що обидві стратегії трансферного навчання дали вищу точність класифікації порівняно з іншими двома підходами. Найкращого результату вдалося досягти при скиданнs FC шару перед донавчанням (F1-оцінка = 0.907, $\sigma = 0.074$). Статистичні тести Вілкоксона підтвердили, що ці покращення є статистично значущими навіть за використання лише кількох повторень для донавчання. Ба більше, трансферне навчання з чотирма повтореннями забезпечило подібну точність, як і навчання з нуля на восьми повтореннях.
Отримані результати свідчать, що донавчання попередньо натренованих моделей може суттєво зменшити зусилля, необхідні для адаптації EMG-систем до нових користувачів, і є практичним та ефективним підходом для створення інтерфейсів, орієнтованих на допоміжне та реабілітаційне застосування.
}
{
поверхнева електроміографія (sEMG), розпізнавання рухів, трансферне навчання, глибоке навчання, згорткові нейронні мережі (CNN), міжіндивідуальна варіативність, узагальнення моделей, стратегії донавчання, персоналізація моделей, міоелектричне керування, ЕМГ-інтерфейс, перехресна перевірка (cross-validation), обробка біомедичних сигналів, реабілітаційні технології, спрощене калібрування
}

\label{radap20NN:LastPage}
\end{multicols}
\newpage
%%%%%%%%%%%%%  END of GENERATED TEXT  %%%%%%%%%%%%\n